Modelling BCG vaccination in the UK

### 2019-12-16 rmds/06-tb-bcg-improv-outcomes.Rmd
updated description of direct effects equations

Bacillus Calmette–Guérin (BCG) primarily reduces the progression from infection to disease, however there is evidence that BCG may provide additional benefits. In this Chapter I aimed to investigate whether there is evidence in routinely-collected surveillance data (see Chapter \@ref(data)) that BCG vaccination impacts outcomes for tuberculosis (TB) cases in England. Any impact on TB outcomes could add additional weight to vaccination policies with wider population coverage, as these policies would have benefits beyond reducing TB incidence rates.

@@

* The code for the analysis contained in this Chapter can be found at: [doi.org/10.5281/zenodo.1213799](doi.org/10.5281/zenodo.1213799)^[Alternatively available from: https://github.com/seabbs/ExploreBCGOnOutcomes]

>>>>

Bacillus Calmette–Guérin (BCG) primarily reduces the progression from infection to disease, however there is evidence that BCG may provide additional benefits. In this chapter I aimed to investigate whether there is evidence in routinely-collected surveillance data (see Chapter \@ref(data)) that BCG vaccination impacts outcomes for tuberculosis (TB) cases in England. Any impact on TB outcomes could add additional weight to vaccination policies with wider population coverage, as these policies would have benefits beyond reducing TB incidence rates.

@@

* The code for the analysis contained in this chapter can be found at: [doi.org/10.5281/zenodo.1213799](doi.org/10.5281/zenodo.1213799)^[Alternatively available from: https://github.com/seabbs/ExploreBCGOnOutcomes]

### 2019-12-16 rmds/07-direct-effect-bcg-change.Rmd
updated description of direct effects equations

In 2005, England changed from universal Bacillus Calmette–Guérin (BCG) vaccination of school-age children to targeted BCG vaccination of high-risk children at birth. In this Chapter I aimed to assess the effects of this change in vaccination policy on the populations targeted by each vaccination scheme.

I combined notification data from the Enhanced TB Surveillance (ETS) system, with demographic data from the Labour Force Survey (LFS) to construct retrospective cohorts of individuals in England relevant to both the universal, and targeted vaccination programmes between Jan 1, 2000 and Dec 31, 2010. For each cohort, I estimated incidence rates over a 5 year follow-up period and used Poisson and negative binomial regression models in order to estimate the impact of the change in policy on TB. This work was adapted from a preprint^[Preprint: https://doi.org/10.1101/567511] supervised by Hannah Christensen and Ellen Brooks-Pollock. Nicky Welton provided guidance on the statistical methods used.

@@

\[ C^i_P = C_0 (1 - I^i),\ \text{Where}\ i = e,\ l,\ u.\]

Where $C^i_P$ is the predicted number of cases prevented using the median ($e$), 2.5% bound ($l$) and 97.5% bound ($u$) of the IRR estimate ($I^i$). For the cohorts relevant to the targeted high-risk neonatal scheme I used a related equation, adjusting for the fact that the populations were exposed to the scheme and I therefore had to first estimate the number of cases that would have been observed had the scheme not been implemented. After simplification this results in the following equation,

\[ C^i_P =  C_0\left(\frac{1}{I^i } - 1\right),\ \text{Where}\ i = e,\ l,\ u.\]

@@

* The code for the analysis contained in this Chapter can be found at: [doi.org/10.5281/zenodo.2583056](doi.org/10.5281/zenodo.2583056)^[Alternatively code link: https://github.com/seabbs/DirectEffBCGPolicyChange]

>>>>

In 2005, England changed from universal Bacillus Calmette–Guérin (BCG) vaccination of school-age children to targeted BCG vaccination of high-risk children at birth. In this chapter I aimed to assess the effects of this change in vaccination policy on the populations targeted by each vaccination scheme.

I combined notification data from the Enhanced TB Surveillance (ETS) system, with demographic data from the Labour Force Survey (LFS) to construct retrospective cohorts of individuals in England relevant to both the universal, and targeted vaccination programmes between Jan 1, 2000 and Dec 31, 2010. For each cohort, I estimated incidence rates over a 5 year follow-up period and used Poisson and negative binomial regression models in order to estimate the impact of the change in policy on TB. This work was adapted from [@Abbott567511]^[Paper: https://doi.org/10.2807/1560-7917.ES.2019.24.49.1900220 Preprint: https://doi.org/10.1101/567511] supervised by Hannah Christensen and Ellen Brooks-Pollock.[@Abbott567511] Nicky Welton provided guidance on the statistical methods used.

@@

\begin{equation}
C^i_P = C_0 (1 - I^i),\ \text{Where}\ i = e,\ l,\ u.
\end{equation}

Where $C^i_P$ is the predicted number of cases prevented using the median ($e$), 2.5% bound ($l$) and 97.5% bound ($u$) of the IRR estimate ($I^i$). For the cohorts relevant to the targeted high-risk neonatal scheme I used a related equation,

\begin{equation}
C^i_P = C_{UE}(1 - I^i),\ \text{Where}\ i = e,\ l,\ u.
\end{equation}

where $C_{UE}$ is the number of cases observed assuming that the cohort was not exposed to targeted high-risk neonatal vaccination. As from 2005 onwards this cohort were in fact exposed to this vaccination scheme an additional step was required. This first required calculating the number of cases that would be expected if the cohort had not been exposed to the scheme,

\begin{equation}
C_{UE} = \frac{C_0}{I^i}
\end{equation}

and then combining this with the previous equation so that $C^i_P$ can be estimated with observed data ($C_0$).

\begin{equation}
C^i_P =  \frac{C_0(1 - I^i)}{I^i},\ \text{Where}\ i = e,\ l,\ u.
\end{equation}

@@

* The code for the analysis contained in this chapter can be found at: [doi.org/10.5281/zenodo.2583056](doi.org/10.5281/zenodo.2583056)^[Alternatively code link: https://github.com/seabbs/DirectEffBCGPolicyChange]
